Reintroduction of resistant frogs facilitates landscape-scale recovery in the presence of a lethal fungal disease

### 2024-04-18 doc/manuscript/translocation_bioarxiv.tex
Render to latex using revised template

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{}
\authordeclaration{}
\equalauthors{\textsuperscript{1}M.W. contributed equally to this work with A.B.}
\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: roland.knapp@ucsb.edu}

% At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
\keywords{evolution | resistance | amphibians | population recovery | reintroduction | disease | \emph{Batrachochytrium dendrobatidis}}

@@

In the current study, we had three primary objectives. First, determine
whether the reintroduction of resistant MYL frogs, obtained from
populations recovering from Bd-caused declines, allows the
reestablishment of populations despite ongoing disease
(Figure~\ref{fig-recovery-model}). We accomplished this using a 15-year
large-scale reintroduction study. Second, extend our inferences of
population recovery beyond the temporal extent of our reintroduction
study to estimate the probability of persistence for the reintroduced
populations over a multi-decadal period
(Figure~\ref{fig-recovery-model}). We did this using a stage-structured
matrix model. Third, given the importance of resistance for frog
survival, population establishment, and long-term viability (this
study), determine whether MYL frogs in recovering populations show
genomic patterns consistent with selection for resistance
(Figure~\ref{fig-recovery-model}). We achieved this with a comparative
study using exome capture methods applied to naive and recovering
populations.

@@

It is important to note that our genomic analysis is necessarily limited
by the small number of naive populations remaining on the landscape
(nearly all of which we sampled as part of the current study) and their
clumped distribution Figure~\ref{fig-allelefrequencies}. These
characteristics of naive populations precluded a replicated analyses
among pairs of geographically proximate naive and recovering
populations. Instead, to obtain adequate power, we conducted a pooled
analysis across population types to identify genomic regions showing
patterns consistent with selection. (See \textbf{Supporting Information
- Frog evolution in response to Bd - Study design} for additional
details regarding the study design.)

As a first step in our pooled analysis, we conducted a principal
component analysis (PCA) of the genomic data to describe the
relationships between sampled populations. We then used two
complementary approaches to identify regions of the genome that differed
between naive and recovering populations (i.e., regions under
selection). First, we used a multivariate linear mixed model to evaluate
associations between population type (i.e., naive versus recovering) and
individual variants, including single nucleotide polymorphisms (SNPs)
and insertions/deletions (INDELS), while accounting for population
structure. Second, we used a splined window analysis to identify larger
genomic regions showing differences between population types in
\emph{F\textsubscript{ST}} and nucleotide diversity (\(\pi_{diff}\) =
\(\pi_{naive}\) - \(\pi_{recovering}\)).

@@

Results from the individual variant and splined window analyses show
that recovering populations show patterns consistent with selection on
multiple regions of the genome. The analysis of individual variants
identified 11 ``outlier'' SNPs (i.e., showing significantly different
allele frequencies between naive versus recovering populations) from 7
distinct genes across 4 contigs (Figure S6B, C
SI). One of the 7 identified genes (LOC108802036) does not have an
associated annotation. For the outlier SNPs, frequency differences
between the naive and recovering populations ranged from 0.41 to 0.86.
Most of these SNPs showed frequency differences in only a subset of the
sampled populations (Figure~\ref{fig-allelefrequencies}A, B), but the
SNP in the RIN3 gene showed consistent differences in frequencies across
all populations (Figure~\ref{fig-allelefrequencies}C). This is
suggestive of parallel selection at this locus across multiple
populations. The other 6 outlier variants showed less consistent
frequency differences across the study populations, but for these we
still found a statistically significant signal of selection in 2 of the
3 genetic clusters (containing populations 5--9;
Figure~\ref{fig-allelefrequencies}, Figure S6A). Therefore, although some outlier variant associations have a more
limited geographic extent than RIN3, they still describe results that
suggest parallel evolutionary changes following Bd exposure.

The splined window analysis identified 33 outlier regions for
\(\pi_{diff}\) and 58 outlier regions for \emph{F\textsubscript{ST}}
(Figure~\ref{fig-spline-manhattan}A, B). Of these, 9 regions were
outliers for both metrics (``shared regions'') and 2 of these shared
regions also contained one or more of the outlier SNPs described above.
A total of 35 annotated genes were found in the 9 shared regions. Given
this large number of genes, here we focus on those with the strongest
signal of selection and/or immune-related functions. The largest
\(\pi_{diff}\), indicative of directional selection, occurred in a 163kb
region on Contig19, 12.9Mb upstream of the RIN3 outlier SNP
(Figure~\ref{fig-spline-manhattan}C). This region contains
approximately 500 SNPs and one annotated gene called
``interferon-induced very large GTPase 1-like'' (GVINP1). Additionally,
a shared outlier region on Contig1 contained two complement factor genes
(C6 and C7). Interestingly, this region had a large negative
\(\pi_{diff}\) value, consistent with balancing selection. Finally, one
shared outlier region on Contig8 contained one outlier SNP (TCF19) and
was within 360kb of another outlier SNP (VARS)
(Figure~\ref{fig-spline-manhattan}D, Figure S7).
This region (854kb from the beginning of the outlier window to the VARS
SNP) contained a total of 8 annotated genes. In \emph{Xenopus}, five of
these genes occur in the extended major histocompatibility complex (MHC)
Class I region (FLOT1, TUBB, MDC1, CCHCR1, TCF19) and three occur in the
extended MHC Class III region (HSP70, LSM2, VARS) \citep{ohta2006}.
Therefore, this region under selection is part of the extended MHC Class
I and III complex and shows synteny with other amphibian genomes.

Although the joint processes of Bd-caused population declines and
selection in response to Bd exposure could affect genetic diversity of
recovering populations, we found no consistent differences in
individual-level heterozygosity or population-level \(\pi\) between
naive and recovering populations (see \textbf{Supporting Information -
Frog evolution in response to Bd - Genetic diversity} for details).
Thus, despite localized selection in particular regions of the genome,
we did not find evidence for reduced genetic diversity across the genome
in recovering populations. In addition, no gene ontology (GO) biological
functions, molecular functions, or cellular processes were
over-represented in either the outlier variants or the 35 genes located
in the overlapping \emph{F\textsubscript{ST}} and \(\pi_{diff}\) splined
windows (see \textbf{Supporting Information - Frog evolution in response
to Bd} \textbf{- GO analysis} for details).

@@

Disease-induced population declines are decimating global biodiversity
\citep{daszak2000}, but broadly-applicable strategies to recover
affected species are generally lacking \citep[e.g.,][]{garner2016}.
Here, we tested the possibility that populations of resistant
individuals from naturally recovering populations can be used to
reestablish extirpated populations of the endangered MYL frog in the
presence of a highly virulent fungal pathogen (Bd). Our results indicate
(i) the capacity of reintroduced populations to become established and
eventually recover despite ongoing disease, (ii) that the recovering
populations are likely to persist over a 50-year period, (iii) that
there are substantial genomic differences between naive and recovering
MYL frog populations, consistent with evolutionary change in frogs
following Bd exposure, and (iv) that some of the genomic regions under
selection contain genes related to disease resistance. Collectively,
these results (Figure~\ref{fig-recovery-model}) provide a rare example
of amphibian recovery in the presence of Bd, and have important
implications for the conservation and recovery of amphibians and other
taxa worldwide that are endangered by escalating impacts from emerging
infectious diseases. In light of the generally low success rate of
amphibian reintroduction efforts \citep{dodd2005}, our success in
reestablishing MYL frog populations via translocation of resistant
individuals is striking, and even more so given that MYL frogs were
driven to near-extinction by Bd.

In the following discussion, we follow the sequence of frog recovery
described in Figure~\ref{fig-recovery-model} to structure our key
points. Previous field studies in MYL frogs show that frog-Bd dynamics
and frog survival in the presence of Bd are fundamentally different
between naive and recovering populations. Following the arrival and
establishment of Bd in previously-naive populations, adult frogs develop
high Bd loads that lead to mass die-offs \citep{vredenburg2010}. In
contrast, in recovering populations adult frogs typically have
low-to-moderate and relatively constant Bd loads and mass die-offs are
not observed \citep[see also Figure S2]{briggs2010, knapp2011}. The differences in Bd load of frogs from
naive and recovering populations are also observed in controlled
laboratory studies \citep[see Figure 4 in][]{knapp2016}, and clearly
indicate that frogs from recovering populations exhibit resistance
against Bd infection. This resistance could in theory be due to several
factors, including natural selection for more resistant genotypes,
acquired immunity, and/or inherent between-population differences that
pre-date Bd exposure, but until now evidence to evaluate the role of
evolution was lacking.

Our genomics study used a pooled design that compared MYL frogs from
naive and recovering populations distributed across a similar
latitudinal range. The study design produced frog genetic clusters that
each contained at least one naive and one recovering population,
allowing us to infer selection while attempting to minimize the
confounding effects of population structure. We identified multiple
specific alleles and genomic regions showing differences between naive
and recovering MYL frog populations, consistent with selection following
Bd exposure. In addition, we did not find a reduction in overall genetic
variation in recovering populations, suggesting that despite localized
selection in the genome, some of these populations likely retain
adequate genetic diversity for long-term persistence.

Importantly, some genomic regions that we identified as possibly under
selection are associated with cellular and immunological mechanisms
known to contribute to disease resistance, including in amphibians
\citep{zamudio2020}. For example, the MHC plays an important role in
immunity. In our study, we identified a region that shows evidence of
selection in recovering populations and contains eight genes associated
with either the MHC Class I or Class III regions. These results
corroborate numerous previous studies linking MHC genes to amphibian
resistance against Bd \citep[e.g.,][]{savage2011, bataille2015}.
Similarly, the region with the strongest indication of directional
selection (as measured by \(\pi_{diff}\)) contains the
interferon-related gene GVINP1. Several previous studies of amphibians
have found this gene to be differentially expressed during Bd infection
\citep[e.g.,][]{grogan2018b, ellison2014} and in populations differing
in Bd susceptibility \citep{grogan2018b}. This gene is also strongly
linked to disease in salmon, explaining a notable 20\% of the resistance
phenotype \citep[i.e., gill damage and amoebic load as heritable host
resistance traits,][]{robledo2020, robledo2018}. We also identified a
region, characterized by high \emph{F\textsubscript{ST}} and low
\(\pi_{diff}\), that contained the complement genes C6 and C7. The
complement system plays an important role in innate immunity
\citep{rieraromo2016}, and our results could indicate that balancing
selection is acting in this region of the genome to favor a diverse set
of alleles, as is known for C6 in humans \citep{soejima2005}. Based on
the analysis of individual outlier variants, the RIN3 gene showed a
consistent pattern of allele frequency differences across all nine of
the frog populations sampled in this study, indicating consistent
selection in populations distributed across a wide geographic area. This
gene is associated with immune response and in \emph{Xenopus} is
expressed during appendage regeneration \citep{fukazawa2009}. Finally,
the outlier variant with the lowest p-value was the uncharacterized gene
LOC108802036. In the genome of another frog species, this gene is
located adjacent to a type I interferon gene (Np-IFNi2) \citep{gan2018},
and together with GVINP1 further suggests the importance of
interferon-related genes in this system. Collectively, the genes
associated with these genomic differences may confer at least some
degree of resistance against Bd infection, an attribute that may be
critically important to population reestablishment and recovery in the
presence of Bd.

@@

In our rapidly changing world, evolution is likely to play in important
role in facilitating the resilience of wildlife populations. Whether the
documented disease resistance in MYL frogs and concurrent recovery of
decimated populations provides an airtight example of evolutionary
rescue will likely always be uncertain (given that can never have a
perfect understanding of the past). Regardless, we provide an example
from the wild that suggests that evolution can produce individuals that
harbor adaptive alleles and allow population recovery in a novel (i.e.,
Bd-positive) environment, and show conclusively that individuals from
these recovering populations can be used to reestablish extirpated
populations and expand the scale of natural recovery
(Figure~\ref{fig-recovery-model}). We expect that similar species
recovery actions will be an essential tool in wildlife conservation in
an era of accelerating global change.

@@

To visualize the genomic relationships of our populations, we conducted
a PCA using the glPCA function in the adegenet R package
\citep{jombart2008}. To detect regions of the genome that differed
between naive and recovering populations, i.e., regions under selection,
we used two approaches: (1) a multivariate linear mixed model to
evaluate individual variants (SNPs and INDELs), and (2) a splined window
analysis to evaluate larger genomic regions. For the variant analysis,
we first used a stringent data filter to include only variants with
\textless{} 5\% missing data (missing for no more than 2 individuals),
and then calculated the likelihood ratio statistic for the resulting set
of 148,307 high quality variants across 127 contigs using GEMMA
\citep{zhou2014}. GEMMA calculates and incorporates a relatedness matrix
for input samples, allowing us to account for relatedness and population
structure when calculating likelihood ratio statistics. We identified
variants showing different allele frequencies between naive versus
recovering populations (``outliers'') using a Bonferroni-corrected
significance level of 0.01. We visualized the results using a Manhattan
plot and qqplot. We developed a more liberal set of outlier variants
using a Bonferroni-corrected significance level of 0.05 and used this
set solely for the gene ontology (GO) analysis (see below and
\textbf{Supporting Information - Frog evolution in response to Bd - GO
analysis;}  Dataset S1, S2).

@@

\acknow{We thank the following for important contributions to this study: E.
Hegeman and A. Lindauer--project and data management, field and
laboratory work; A. Barbella and K. Rose--laboratory work; C. Kamoroff
and numerous summer technicians--field work; staff at Sequoia-Kings
Canyon and Yosemite National Parks, Inyo and Sierra National Forests,
California Department of Fish and Wildlife, U.S. Fish and Wildlife
Service, University of California-Santa Barbara Institutional Animal
Care and Use Committee, and Sierra Nevada Aquatic Research
Laboratory--research permits, logistical support, and/or field
assistance. This project was supported by grants from the National Park
Service (to R.A.K.), Yosemite Conservancy (to R.A.K.), and National
Science Foundation (EF-0723563, to C. Briggs; DEB-1557190, to C. Briggs;
DEB-2133401, to M.Q.W; and DBI-2120084, to C. Richards-Zawacki).}

>>>>

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{All authors designed research; R.A.K., M.Q.W., A.Q.B., M.B.J., T.C.S., A.P.R., and R.L.G. performed research; M.B.J. contributed new analytic tools; R.A.K., M.Q.W., A.Q.B., A.P.R., and E.B.R. analyzed data; and R.A.K., M.Q.W., A.Q.B., M.B.J., T.C.S., and E.B.R. wrote paper.}
\authordeclaration{This project was supported by grants from the National Park Service (to R.A.K.), Yosemite Conservancy (to R.A.K.), and National Science Foundation (EF-0723563, to C. Briggs; DEB-1557190, to C. Briggs; DEB-2133401, to M.Q.W; and DBI-2120084, to C. Richards-Zawacki). The authors declare no additional competing interests.}
\equalauthors{\textsuperscript{1}M.W. contributed equally to this work with A.B.}
\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: roland.knapp@ucsb.edu}

% At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
\keywords{amphibians | disease | evolution | reintroduction | population recovery}

@@

In the current study, we had three primary objectives. First, determine
whether the reintroduction of resistant MYL frogs, obtained from
populations recovering from Bd-caused declines, allows the
reestablishment of populations despite ongoing disease
(Figure~\ref{fig-recovery-model}). We accomplished this using a 15-year
large-scale reintroduction study. Second, extend our inferences of
population recovery beyond the temporal bounds of our reintroduction
study to estimate the probability of persistence for the reintroduced
populations over a multi-decadal period
(Figure~\ref{fig-recovery-model}). We did this using a stage-structured
matrix model. Third, given the importance of resistance for frog
survival, population establishment, and long-term viability (this
study), determine whether MYL frogs in recovering populations show
genomic patterns consistent with selection for resistance
(Figure~\ref{fig-recovery-model}). We achieved this with a comparative
study using exome capture methods applied to naive and recovering
populations.

@@

It is important to note that our genomic analysis is necessarily limited
by the small number of naive populations remaining on the landscape
(nearly all of which we sampled as part of the current study) and their
clumped distribution (Figure~\ref{fig-allelefrequencies}). These
characteristics of naive populations precluded a replicated analyses
among pairs of geographically proximate naive and recovering
populations. Instead, to obtain adequate power, we conducted a pooled
analysis across population types to identify genomic regions showing
patterns consistent with selection. Our comparative approach cannot
provide a direct assessment of the strength or timing of natural
selection, but does allow us to identify specific regions of the genome
that merit further study for their potential contribution to frog
resistance in recovering populations. (See \textbf{Supporting
Information - Frog evolution in response to Bd - Study design} for
additional details regarding the study design.)

As a first step in our pooled analysis, we conducted a principal
component analysis (PCA) of the genomic data to describe the
relationships between sampled populations. We then used two
complementary approaches to identify regions of the genome that differed
between naive and recovering populations (i.e., regions possibly under
selection). First, we used a multivariate linear mixed model to evaluate
associations between population type (i.e., naive versus recovering) and
individual variants, including single nucleotide polymorphisms (SNPs)
and insertions/deletions (INDELS), while accounting for population
structure. Second, we used a splined window analysis to identify larger
genomic regions showing differences between population types in
\emph{F\textsubscript{ST}} and nucleotide diversity (\(\pi_{diff}\) =
\(\pi_{naive}\) - \(\pi_{recovering}\)).

@@

Results from the individual variant and splined window analyses show
that recovering populations show patterns consistent with selection on
multiple regions of the genome. The analysis of individual variants
identified 11 ``outlier'' SNPs (i.e., showing significantly different
allele frequencies between naive versus recovering populations) from 7
distinct genes across 4 contigs (Figure S6B, C
SI). One of the 7 identified genes (LOC108802036) does not have an
associated annotation. For the outlier SNPs, frequency differences
between the naive and recovering populations ranged from 0.41 to 0.86.
Most of these SNPs showed frequency differences in only a subset of the
sampled populations (Figure~\ref{fig-allelefrequencies}A, B), but the
SNP in the RIN3 gene showed consistent differences in frequencies across
all populations (Figure~\ref{fig-allelefrequencies}C). This is
suggestive of parallel selection at this locus across multiple
populations. The other 6 outlier variants showed less consistent
frequency differences across the study populations, but for these we
still found statistically significant differences in 2 of the 3 genetic
clusters (containing populations 5--9;
Figure~\ref{fig-allelefrequencies}, Figure S6A). Therefore, although some outlier variant associations have a more
limited geographic extent than RIN3, they still describe results that
are consistent with parallel evolutionary changes following Bd exposure.

The splined window analysis identified 33 outlier regions for
\(\pi_{diff}\) and 58 outlier regions for \emph{F\textsubscript{ST}}
(Figure~\ref{fig-spline-manhattan}A, B). Of these, 9 regions were
outliers for both metrics (``shared regions'') and 2 of these shared
regions also contained one or more of the outlier SNPs described above.
A total of 35 annotated genes were found in the 9 shared regions. Given
this large number of genes, here we focus on those with the strongest
signal of selection and/or immune-related functions. The largest
\(\pi_{diff}\), suggestive of directional selection, occurred in a 163kb
region on Contig19, 12.9Mb upstream of the RIN3 outlier SNP
(Figure~\ref{fig-spline-manhattan}C). This region contains
approximately 500 SNPs and one annotated gene called
``interferon-induced very large GTPase 1-like'' (GVINP1). Additionally,
a shared outlier region on Contig1 contained two complement factor genes
(C6 and C7). Interestingly, this region had a large negative
\(\pi_{diff}\) value, consistent with balancing selection. Finally, one
shared outlier region on Contig8 contained one outlier SNP (TCF19) and
was within 360kb of another outlier SNP (VARS)
(Figure~\ref{fig-spline-manhattan}D, Figure S7).
This region (854kb from the beginning of the outlier window to the VARS
SNP) contained a total of 8 annotated genes. In \emph{Xenopus}, five of
these genes occur in the extended major histocompatibility complex (MHC)
Class I region (FLOT1, TUBB, MDC1, CCHCR1, TCF19) and three occur in the
extended MHC Class III region (HSP70, LSM2, VARS) \citep{ohta2006}.
Therefore, this region is part of the extended MHC Class I and III
complex and shows synteny with other amphibian genomes.

Although the joint processes of Bd-caused population declines and
selection in response to Bd exposure could affect genetic diversity of
recovering populations, we found no consistent differences in
individual-level heterozygosity or population-level \(\pi\) between
naive and recovering populations (see \textbf{Supporting Information -
Frog evolution in response to Bd - Genetic diversity} for details).
Thus, despite patterns consistent with localized selection in particular
regions of the genome, we did not find evidence for reduced genetic
diversity across the genome in recovering populations. In addition, no
gene ontology (GO) biological functions, molecular functions, or
cellular processes were over-represented in either the outlier variants
or the 35 genes located in the overlapping \emph{F\textsubscript{ST}}
and \(\pi_{diff}\) splined windows (see \textbf{Supporting Information -
Frog evolution in response to Bd} \textbf{- GO analysis} for details).

@@

Disease-induced population declines are decimating global biodiversity
\citep{daszak2000}, but broadly-applicable strategies to recover
affected species are generally lacking \citep[e.g.,][]{garner2016}.
Here, we tested the possibility that populations of resistant
individuals from naturally recovering populations can be used to
reestablish extirpated populations of the endangered MYL frog in the
presence of a highly virulent fungal pathogen (Bd). Our results indicate
(i) the capacity of reintroduced populations to become established and
eventually recover despite ongoing disease, (ii) that 50\% of the
recovering populations are likely to persist over a 50-year period,
(iii) that there are genomic differences between naive and recovering
MYL frog populations, consistent with evolutionary change in frogs
following Bd exposure, and (iv) that some of the genomic differences are
in regions that contain genes related to disease resistance.
Collectively, these results (Figure~\ref{fig-recovery-model}) provide a
rare example of amphibian recovery in the presence of Bd, and have
important implications for the conservation and recovery of amphibians
and other taxa worldwide that are endangered by escalating impacts from
emerging infectious diseases. In light of the generally low success rate
of amphibian reintroduction efforts \citep{dodd2005}, our success in
reestablishing MYL frog populations via translocation of resistant
individuals is striking, and even more so given that MYL frogs were
driven to near-extinction by Bd.

In the following discussion, we follow the sequence of frog recovery
described in Figure~\ref{fig-recovery-model} to structure our key
points. Previous field studies in MYL frogs show that frog-Bd dynamics
and frog survival in the presence of Bd are fundamentally different
between naive and recovering populations. Following the arrival and
establishment of Bd in previously-naive populations, adult frogs develop
high Bd loads that lead to mass die-offs \citep{vredenburg2010}. In
contrast, in recovering populations adult frogs typically have
low-to-moderate and relatively constant Bd loads and mass die-offs are
not observed \citep[see also Figure S2]{briggs2010, knapp2011}. The differences in Bd load of frogs from
naive and recovering populations are also observed in controlled
laboratory studies \citep[see Figure 4 in][]{knapp2016}, and clearly
indicate that frogs from recovering populations exhibit resistance
against Bd infection. This resistance could be due at least in part to
natural selection for more resistant genotypes, but until now evidence
to evaluate the possible role of evolution was lacking.

Our genomics study used a pooled design that compared MYL frogs from
naive and recovering populations distributed across a similar
latitudinal range. The study design produced frog genetic clusters that
each contained at least one naive and one recovering population,
allowing us to infer selection while attempting to minimize the
confounding effects of population structure. We identified multiple
specific alleles and genomic regions showing differences between naive
and recovering MYL frog populations, consistent with selection following
Bd exposure. In addition, we did not find a reduction in overall genetic
variation in recovering populations, suggesting that despite possible
localized selection in the genome, some of these populations likely
retain adequate genetic diversity for long-term persistence.

Importantly, some genomic regions that we identified as possibly under
selection are associated with cellular and immunological mechanisms
known to contribute to disease resistance, including in amphibians
\citep{zamudio2020}. For example, the MHC plays an important role in
immunity. In our study, we identified a region that shows evidence of
selection in recovering populations and contains eight genes associated
with either the MHC Class I or Class III regions. These results
corroborate numerous previous studies linking MHC genes to amphibian
resistance against Bd \citep[e.g.,][]{savage2011, bataille2015}.
Similarly, the region with the strongest pattern of possible directional
selection (as measured by \(\pi_{diff}\)) contains the
interferon-related gene GVINP1. Several previous studies of amphibians
have found this gene to be differentially expressed during Bd infection
\citep[e.g.,][]{grogan2018b, ellison2014} and in populations differing
in Bd susceptibility \citep{grogan2018b}. This gene is also strongly
linked to disease in salmon, explaining a notable 20\% of the resistance
phenotype \citep[i.e., gill damage and amoebic load as heritable host
resistance traits,][]{robledo2020, robledo2018}. We also identified a
region, characterized by high \emph{F\textsubscript{ST}} and low
\(\pi_{diff}\), that contained the complement genes C6 and C7. The
complement system plays an important role in innate immunity
\citep{rieraromo2016}, and our results could indicate that balancing
selection is acting in this region of the genome to favor a diverse set
of alleles, as is known for C6 in humans \citep{soejima2005}. Based on
the analysis of individual outlier variants, the RIN3 gene showed a
consistent pattern of allele frequency differences across all nine of
the frog populations sampled in this study, suggesting possible
consistent selection in populations distributed across a wide geographic
area. This gene is associated with immune response and in \emph{Xenopus}
is expressed during appendage regeneration \citep{fukazawa2009}.
Finally, the outlier variant with the lowest p-value was the
uncharacterized gene LOC108802036. In the genome of another frog
species, this gene is located adjacent to a type I interferon gene
(Np-IFNi2) \citep{gan2018}, and together with GVINP1 further suggests
the importance of interferon-related genes in this system. Collectively,
the genes associated with these genomic differences may confer at least
some degree of resistance against Bd infection, an attribute that may be
critically important to population reestablishment and recovery in the
presence of Bd.

@@

In our rapidly changing world, evolution is likely to play in important
role in facilitating the resilience of wildlife populations. Whether the
documented disease resistance in MYL frogs and concurrent recovery of
decimated populations provides an airtight example of evolutionary
rescue will likely always be uncertain (given that we can never have a
perfect understanding of the past). Regardless, we provide an example
from the wild that suggests that evolution might produce individuals
that harbor adaptive alleles and allow population recovery in a novel
(i.e., Bd-positive) environment, and show conclusively that individuals
from these recovering populations can be used to reestablish extirpated
populations and expand the scale of natural recovery
(Figure~\ref{fig-recovery-model}). We expect that similar species
recovery actions will be an essential tool in wildlife conservation in
an era of accelerating global change.

@@

To visualize the genomic relationships of our populations, we conducted
a PCA using the glPCA function in the adegenet R package
\citep{jombart2008}. To detect regions of the genome that differed
between naive and recovering populations, we used two approaches: (1) a
multivariate linear mixed model to evaluate individual variants (SNPs
and INDELs), and (2) a splined window analysis to evaluate larger
genomic regions. For the variant analysis, we first used a stringent
data filter to include only variants with \textless{} 5\% missing data
(missing for no more than 2 individuals), and then calculated the
likelihood ratio statistic for the resulting set of 148,307 high quality
variants across 127 contigs using GEMMA \citep{zhou2014}. GEMMA
calculates and incorporates a relatedness matrix for input samples,
allowing us to account for relatedness and population structure when
calculating likelihood ratio statistics. We identified variants showing
different allele frequencies between naive versus recovering populations
(``outliers'') using a Bonferroni-corrected significance level of 0.01.
We visualized the results using a Manhattan plot and qqplot. We
developed a more liberal set of outlier variants using a
Bonferroni-corrected significance level of 0.05 and used this set solely
for the gene ontology (GO) analysis (see below and \textbf{Supporting
Information - Frog evolution in response to Bd - GO analysis;}Dataset
S1, S2).

@@

\acknow{We thank the following for important contributions to this study: E.
Hegeman and A. Lindauer (project and data management, field and
laboratory work); A. Barbella and K. Rose-(laboratory work); C. Kamoroff
and numerous summer technicians (field work); staff at Sequoia-Kings
Canyon and Yosemite National Parks, Inyo and Sierra National Forests,
California Department of Fish and Wildlife, U.S. Fish and Wildlife
Service, University of California-Santa Barbara Institutional Animal
Care and Use Committee, and Sierra Nevada Aquatic Research Laboratory
(research permits, logistical support, and/or field assistance). This
project was supported by grants from the National Park Service (to
R.A.K.), Yosemite Conservancy (to R.A.K.), and National Science
Foundation (EF-0723563, to C. Briggs; DEB-1557190, to C. Briggs;
DEB-2133401, to M.Q.W; and DBI-2120084, to C. Richards-Zawacki).}
